\documentclass[runningheads,a4paper]{llncs}
%\documentclass[letter,11pt]{article}
%geometry -> one inch marge.
\usepackage{./sty/mystyle}

\usepackage{url}
\urldef{\mailsa}\path|mateusz.lacki@biol.uw.edu.pl|
\urldef{\mailsb}\path|aniag@mimuw.edu.pl|

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

\begin{document}
\mainmatter  

\title{Law of Localised Fine Structure\thanks{
	This research was partially supported by Polish National Science Center grant $\text{n}^\text{o}$ 2011/01/B/NZ2/00864.
}}

\subtitle{\textit{with application in mass spectrometry proteomic studies.}}

\titlerunning{Mass Spec Fine Structure Distribution}

\author{Mateusz Krzysztof \L\k{a}cki
\and Anna Gambin}

\authorrunning{Fine Structure Distribution in Mass Spectrometry}


% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{
  Faculty of Mathematics, Informatics and Mechanics\\ University of Warsaw, Banacha 2, 02-097 Warszawa, Poland\\
  \mailsa\\
  \mailsb
  % \url{http://bioputer.mimuw.edu.pl}
}

% NB: a more complex sample for affiliations and the mapping to the
% corresponding authors can be found in the file "llncs.dem"
% (search for the string "\mainmatter" where a contribution starts).
% "llncs.dem" accompanies the document class "llncs.cls".

\toctitle{Mass Spec Fine Structure Distribution}
\tocauthor{Mateusz \L\k{a}cki}
\maketitle

\begin{abstract}
	This paper presenst a brand new methodology to deal with isotopic fine structure calculations. Huge improvents in problem's tractability are gained through the use of Poisson approximation to the standard model of isotopic species distribution. The main results are twofolds and can be classified on different levels of the peak aggregation hierarchy. 

	Firstly, they can be used to perform efficient calculations on the most fine level of isotopic distribution. 

	Secondly, calculations on the equatransneutronic clusters can be performed without significant efficiency loss with respect to algorithms based on more aggregated isotopic distribution. 

	our approach can be used to perform calculations 

	Our results can be implemented on two different levels of aggregation .

	on different levels of the peak aggregation hierarchy. 

	Hierarchi	

The distribution of isotopic envelope is approximated indirectly by Poisson laws. We use the devised approximations to design alternative algorithms for calculating the isotopic fine structure with significant coverage. 
\keywords{Isotopic Fine Structure, Poisson Approximation, Stable Isotopes, Avergine Model.}
\end{abstract}

% \listoftodos

\input{./tex/intro}
\input{./tex/approximations}
\input{./tex/algorithms}
\input{./tex/discussionsAndConclusions}


\input{./tex/acknowledgments}
  \bibliographystyle{./bst/splncs03.bst}
  \bibliography{./bib/spectrometry,./bib/probability,./bib/topology,./bib/generalMath,./bib/computerSciences}

  % no \newpage when using \include
\include{./tex/tables}
\include{./tex/figures}
\include{./tex/appendix}


\end{document}
